Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies

### 2018-11-23 man/figures/cumulative_prec.Rmd
fix error size figures

---
output: 
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, dpi = 200)
```

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")
library("grid")
library("gridExtra")
library("patchwork")
library("ggcorrplot")
library("stringr")
library("pander")
library("lemon")
```

```{r, echo=FALSE}
show_text <- FALSE
```

##### 
**Appendix S1.** Temporal evolution of cumulative precipitation (hydrological year) during the period 1950-2017. Points represent mean and errorbars standard error. *Black* line indicates mean for all period. *Red* lines represent -1 and -2 standard deviation (*dotted* and *dashed* lines respectively). *Blue* lines represent +1 and +2 standard deviation (*dotted* and *dashed* lines respectively). Years with average values below -1SD are labelled. Data from 28 meteorological stations distributed around Sierra Nevada area (from National Spanish Meteorological Services, AEMET). ***Inset plot***: cumulative precipitation during the hydrological years 2004-2005 (*blue line*) and 2011-2012 (*red line*). The boxplot representing the average from 1950-2015 period. Data from meteorological station Granada, Base Aérea.

```{r}
ltdf <- read.table(file=here::here("/data_raw/meteo", "long_term_data_rediam_data.txt"), header=T, sep=';', dec=',')

@@

cp <- ggplot(df, aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot(aes(x=as.factor(hmonth))) +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(st5514_pro, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(st5514_pro, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1, size=7),
        axis.text.y = element_text(size=7),
        axis.title.y = element_text(size = 7)) + 
  ylab('Cumulative Precipitation (mm)')  + xlab('') + 
  # ggtitle('Granada Base Aerea, 1950-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug")) +
  annotate("text", label = "2004-2005", x = 2, y = 500, size = 2, colour = "blue") + 
  annotate("text", label = "2011-2012", x = 2, y = 550, size = 2, colour = "red") + 
  annotate("text", label = "1950-2015", x = 3, y = 600, size = 2, colour = "black") +
  theme(panel.grid = element_blank())
```

```{r}
prec_files <- list.files(here::here('data_raw/meteo/prec_subclima/'), pattern="*.csv")

@@

meandf <- mean(presn$csum_mean)
sddf <- sd(presn$csum_mean)

prec_hydrol_sn <- ggplot(presn, aes(x=hyear, y=csum_mean)) + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
  geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, colour='gray') +
  geom_point(size=2) + theme_bw() + 
  geom_line(size=.3) + 
  scale_x_continuous(breaks=seq(1950,2020, by=5)) +
  scale_y_continuous(breaks=seq(0,1500, by=250)) +
  geom_text_repel(
    data = subset(presn, csum_mean < (meandf - 1*sddf)),
    aes(label = hyear), 
    size = 3) +
  theme(panel.grid.minor = element_blank()) +
  ylab('Cumulative Precipitation (hydrological year)') + xlab('year')

prec_hydrol_sn + annotation_custom(ggplotGrob(cp), xmin = 1990, xmax = 2020, ymin = 1050)
```

>>>>

---
output: 
  word_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("ggrepel")
library("scales")
library("grid")
library("gridExtra")
library("patchwork")
library("ggcorrplot")
library("stringr")
library("pander")
library("lemon")
```

#####

```{r data}
ltdf <- read.table(file=here::here("/data_raw/meteo", "long_term_data_rediam_data.txt"), header=T, sep=';', dec=',')

@@

```

```{r plot_cummulatedPrec}

size_line <- .25

cp <- ggplot(df, aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot(aes(x=as.factor(hmonth)), size = 0.25, outlier.size = 0.5) +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black', size = size_line) + 
  geom_line(data=subset(st5514_pro, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue', size = size_line) +
  geom_line(data=subset(st5514_pro, hyear_f == '2011-2012'), 
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red', size = size_line) +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5, size=7),
        axis.text.y = element_text(size=7),
        axis.title.y = element_text(size = 7)) + 
  ylab("Cumulative \n Precipitation (mm)") + xlab('') + 
  # ggtitle('Granada Base Aerea, 1950-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug")) +
  annotate("text", label = "2004-2005", x = 2, y = 420, size = 2, colour = "blue") + 
  annotate("text", label = "2011-2012", x = 2, y = 520, size = 2, colour = "red") + 
  annotate("text", label = "1950-2015", x = 2, y = 620, size = 2, colour = "black") +
  theme(panel.grid = element_blank(), 
        axis.title.y = element_text(hjust = 0.5),
        plot.background = element_blank())
```

```{r data_prec}
prec_files <- list.files(here::here('data_raw/meteo/prec_subclima/'), pattern="*.csv")

@@

meandf <- mean(presn$csum_mean)
sddf <- sd(presn$csum_mean)
```

```{r}

yeah <- presn %>% filter(csum_mean < (meandf - 1*sddf)) %>% 
  mutate(pos = csum_mean - 50) %>%  as.data.frame()

prec_hydrol_sn <- ggplot(presn, aes(x=hyear, y=csum_mean)) + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
  geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, colour='gray') +
  geom_point(size=2) + theme_bw() + 
  geom_line(size=.25, colour = 'black') + 
  scale_x_continuous(breaks=seq(1950,2015, by=5)) +
  scale_y_continuous(breaks=seq(0,1500, by=250)) + 
  geom_text_repel(
    data = yeah,
    aes(label = hyear), 
    size = 2.5, 
    # position = position_dodge(1), 
    alpha = .5,
    colour = "black",
    segment.size = 0.1, 
    segment.alpha = 0.5,
    point.padding=unit(.7,'lines'),
    nudge_y = -yeah$csum_mean,
    direction = 'x'
    ) +
  theme(panel.grid = element_blank(),
        axis.title = element_text(size = 9),
        axis.text.y = element_text(size = 8)) +
  ylab('Cumulative Precipitation (mm)') + xlab('Year') +
  expand_limits(y=c(0,2000)) 
```

```{r}
p <- prec_hydrol_sn + annotation_custom(ggplotGrob(cp), xmin = 1990, xmax = Inf, ymin=1050)
```

**Appendix S1.** Temporal evolution of cumulative precipitation (hydrological year) during the period 1950-2017. Points represent mean and errorbars standard error. *Black* line indicates mean for all period. *Red* lines represent -1 and -2 standard deviation (*dotted* and *dashed* lines respectively). *Blue* lines represent +1 and +2 standard deviation (*dotted* and *dashed* lines respectively). Years with average values below -1SD are labelled. Data from 28 meteorological stations distributed around Sierra Nevada area (from National Spanish Meteorological Services, AEMET). ***Inset plot***: cumulative precipitation during the hydrological years 2004-2005 (*blue line*) and 2011-2012 (*red line*). The boxplot representing the average from 1950-2015 period. Data from meteorological station Granada, Base Aérea.

```{r plot_out, fig.width=7, dpi = 300, out.width = 0.8} 
p
```

### 2018-11-23 man/figures/growth_changes.Rmd
fix error size figures

```{r plot_join }
gg <- psj + pcaH + pcaL + 
  plot_layout(ncol=1) & 
  # theme(plot.tag = element_text(size = 12, face = 2)) & 
  theme(plot.margin = margin(t=-1,r=3,b=-1,l=3, "pt"))
```

```{r out_plot, fig.height=5, dpi = 300, out.width = 0.8}
gg

>>>>

```{r plot_join}
gg <- psj + pcaH + pcaL + 
  plot_layout(ncol=1) & 
  # theme(plot.tag = element_text(size = 12, face = 2)) & 
  theme(plot.margin = margin(t=-1,r=3,b=-1,l=3, "pt"))
```

```{r plot_out, fig.height=5, dpi = 300, out.width = 0.8}
gg

### 2018-11-23 man/figures/resilience_metrics.Rmd
fix error size figures

---
output: 
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, dpi = 200)
```

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")

library("grid")
library("gridExtra")
library("patchwork")
library("ggcorrplot")
library("stringr")
library("pander")
library("lemon")
```

```{r, echo=FALSE}
show_text <- FALSE
```

##### 
**Figure 3.** Response *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (EVI; left-plots) and tree radial growth (BAI; right-plots) for the years 2005 and 2012. For EVI we compared northern populations (*black fill circle*) with southern ones (*blue empty circle*). For BAI we compared northern population (San Juan, SJ; *black triangle*) with southerns populations: Cáñar-High (CA-High; *blue empty squares*) and Cáñar-Low (CA-Low; *blue fill squares*). Different letters above error bars indicate significant *post hoc* differences between groups (see material and methods).

```{r, eval=FALSE, echo=FALSE}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

@@

theme_bai <- list(
  scale_color_manual(values=c('black','blue','blue')), 
  scale_shape_manual(values=c(19, 22, 15)),
  scale_fill_manual(values=c('black','white','blue')),  
  scale_linetype_manual(values=c("solid", "solid", 'solid'))
)
```

```{r, eval=FALSE, echo=FALSE}
# EVI_drought 
res_evi_drought <-
  ggplot(mhuber_filter, aes_disturb) + 
  theme_res + theme_evi + theme(legend.position='top') + 
  ggtitle("a) EVI")

@@

```{r, eval=FALSE, fig.height=7}
res_evi_drought  + res_bai_drought + plot_layout(ncol=1)
```

```{r, fig.height=9}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

aux <- mhuber_filter %>% mutate(vari = '0_EVI')
aux2 <- mhuberbai_filter %>% mutate(vari = '1_BAI')
m <- bind_rows(aux2, aux)

pd <- position_dodge(width = 0.6)

themeR <- list(
  geom_errorbar(width=.1, position = pd),
  geom_point(position = pd),
  theme_bw(),
  facet_wrap(~vari, labeller = labeller(vari = variables), scales="free"),
  xlab(''), 
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(
      colour = "black",
      fill = "white"),
    strip.text = element_text(face="bold", size = 14),
    axis.title.y = element_text(face="bold", size = 14),
    plot.title = element_text(hjust = -0.05, size = 16, margin = margin(t=-5, b=1))),
  scale_shape_manual(labels = item_site, guide = guide_legend(title = "Site"),
    values=c("0_N" = 19, "1_S" = 21, "3_CA-High" = 22, "4_CA-Low" = 15, "2_SJ" = 17)), 
  scale_color_manual(labels = item_site, guide = tl,
    values=c("0_N" = 'black', "1_S" = "blue", "3_CA-High" = 'blue', "4_CA-Low" = 'blue', "2_SJ" = 'black')),
  scale_fill_manual(labels = item_site, guide = tl,
    values=c("0_N" = 'black', "1_S" = "white", "3_CA-High" = 'white', "4_CA-Low" = 'blue', "2_SJ" = 'black')),
  scale_linetype_manual(labels = item_site, guide = tl,
                        values=c("solid","solid","solid","solid", "solid")),
  scale_size_manual(labels = item_site, guide = tl,
    values=c("0_N" = "2", "1_S" = "2", "3_CA-High" = 2.5, "4_CA-Low" = 2.5, "2_SJ" = 2.5)),

rt <- m %>% filter(var == "rt") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resistance") 
# + geom_text(aes(label=Letter), position=position_jitterdodge(jitter.width = .2, dodge.width = 0.6), show.legend = FALSE)

rc <- m %>% filter(var == "rc") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Recovery")
# + geom_text(aes(label=Letter, y = 1.65), position=pd, show.legend = FALSE)

rs <- m %>% filter(var == "rs") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resilience") + xlab('Drought event') 
# + geom_text(aes(label=Letter, y = 1.2), position=pd, show.legend = FALSE) 
```

```{r, fig.height=6}
grid_arrange_shared_legend(rt, rc, rs,ncol = 1, nrow = 3, position='top') 
```

```{r, eval=FALSE}

# BAI_site
res_bai_site <-
  ggplot(mhuberbai_filter, aes_site) + 
  theme_res + theme_de + theme(legend.position='bottom') + 
  xlab('Site') + 
  ggtitle("b) BAI") 
```

```{r, eval=FALSE}
res_evi_site  + res_bai_site + plot_layout(ncol=1)

**Figure .** Response of *Q. pyrenaica* forests to 2005 (*black*) and 2012 (*red*) drought events in terms of resistance, recovery and resilience of greenness (EVI; a) and tree radial growth (BAI; b) by site. Different letters above bars indicate significant post hoc differences between groups (see material and methods). 
```

>>>>

---
output: 
  word_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("ggrepel")
library("grid")
library("gridExtra")
library("patchwork")
library("lemon")
library("bookdown")
```

#####

```{r data}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

# NO rrs
mhuber_filter <- mhuber %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))
mhuberbai_filter <- mhuberbai %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))

aux <- mhuber_filter %>% mutate(vari = '0_EVI')
aux2 <- mhuberbai_filter %>% mutate(vari = '1_BAI')
m <- bind_rows(aux2, aux)
```

```{r plot}
pd <- position_dodge(width = 0.6)

themeR <- list(
  geom_errorbar(width=.1, position = pd),
  geom_point(position = pd),
  theme_bw(),
  facet_wrap(~vari, labeller = labeller(vari = variables), scales="free"),
  xlab(''), 
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(
      colour = "black",
      fill = "white"),
    strip.text = element_text(face="bold", size = 10),
    axis.title.y = element_text(face="bold", size = 10),
    plot.title = element_text(hjust = -0.05, size = 10, margin = margin(t=-5, b=1)),
    axis.text = element_text(size = 9),
    panel.spacing.x = unit(1, "lines")),
  scale_shape_manual(labels = item_site, guide = guide_legend(title = "Site"),
    values=c("0_N" = 19, "1_S" = 21, "3_CA-High" = 22, "4_CA-Low" = 15, "2_SJ" = 17)), 
  scale_color_manual(labels = item_site, guide = tl,
    values=c("0_N" = 'black', "1_S" = "blue", "3_CA-High" = 'blue', "4_CA-Low" = 'blue', "2_SJ" = 'black')),
  scale_fill_manual(labels = item_site, guide = tl,
    values=c("0_N" = 'black', "1_S" = "white", "3_CA-High" = 'white', "4_CA-Low" = 'blue', "2_SJ" = 'black')),
  scale_linetype_manual(labels = item_site, guide = tl,
                        values=c("solid","solid","solid","solid", "solid")),
  scale_size_manual(labels = item_site, guide = tl,
    values=c("0_N" = "2", "1_S" = "2", "3_CA-High" = 2.5, "4_CA-Low" = 2.5, "2_SJ" = 2.5)),

rt <- m %>% filter(var == "rt") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resistance") +
  theme(plot.margin = margin(r=0, l=0, t=0, b=-2, "pt"))
# + geom_text(aes(label=Letter), position=position_jitterdodge(jitter.width = .2, dodge.width = 0.6), show.legend = FALSE)

rc <- m %>% filter(var == "rc") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Recovery") +
  theme(plot.margin = margin(r=0, l=0, t=0, b=-2, "pt"))
# + geom_text(aes(label=Letter, y = 1.65), position=pd, show.legend = FALSE)

rs <- m %>% filter(var == "rs") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resilience") + xlab('Drought event') +
    theme(plot.margin = margin(r=0, l=0, t=0, b=-2, "pt"))
# + geom_text(aes(label=Letter, y = 1.2), position=pd, show.legend = FALSE) 
```

**Figure 7.** Response *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (EVI; left-plots) and tree radial growth (BAI; right-plots) for the years 2005 and 2012. For EVI we compared northern populations (*black fill circle*) with southern ones (*blue empty circle*). For BAI we compared northern population (San Juan, SJ; *black triangle*) with southerns populations: Cáñar-High (CA-High; *blue empty squares*) and Cáñar-Low (CA-Low; *blue fill squares*). Different letters above error bars indicate significant *post hoc* differences between groups (see material and methods).

```{r plot_out, fig.height=6.5, dpi = 300, out.width = 0.8} 
grid_arrange_shared_legend(rt, rc, rs, ncol = 1, nrow = 3, position ='top')

```{r old1, eval=FALSE}

# BAI_site
res_bai_site <-
  ggplot(mhuberbai_filter, aes_site) + 
  theme_res + theme_de + theme(legend.position='bottom') + 
  xlab('Site') + 
  ggtitle("b) BAI")

res_evi_site  + res_bai_site + plot_layout(ncol=1)

# **Figure .** Response of *Q. pyrenaica* forests to 2005 (*black*) and 2012 (*red*) drought events in terms of resistance, recovery and resilience of greenness (EVI; a) and tree radial growth (BAI; b) by site. Different letters above bars indicate significant post hoc differences between groups (see material and methods). 
```

```{r old2, eval=FALSE, echo=FALSE, fig.height=7}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

@@

theme_bai <- list(
  scale_color_manual(values=c('black','blue','blue')), 
  scale_shape_manual(values=c(19, 22, 15)),
  scale_fill_manual(values=c('black','white','blue')),  
  scale_linetype_manual(values=c("solid", "solid", 'solid'))
)

# EVI_drought 
res_evi_drought <-
  ggplot(mhuber_filter, aes_disturb) + 
  theme_res + theme_evi + theme(legend.position='top') + 
  ggtitle("a) EVI")

@@

res_evi_drought  + res_bai_drought + plot_layout(ncol=1)
```

### 2018-11-23 man/figures/ring_climate.Rmd
fix error size figures

---
output: 
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, dpi = 200)
```

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")

@@

```{r, echo=FALSE}
show_text <- FALSE
```

##### 
**Figure 6.** Correlation coefficients obtained by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data: precipitation and 6-month SPEI (a), minimun (b) and maximun (c) temperatures. *green* bars: northern site (SJ); *light blue* bars: low-elevation southern site (CA-Low); and *dark blue* bars: high-elevation shouthern site (CA-High). Asteriks indicate significant ($P < 0.05$) correlation coefficients.

```{r}
rtmax <- read.csv(here("data/rwi_climate", "rtmax_selected.csv"), header = TRUE, sep = ',') %>% mutate(v_name = 'T max')
rtmin <- read.csv(here("data/rwi_climate", "rtmin_selected.csv"), header = TRUE, sep = ',') %>% mutate(v_name = 'T min')
rsequia <- read.csv(here("data/rwi_climate", "rsequia_selected.csv"), header = TRUE, sep = ',') %>% mutate(v_name ='SPEI')
rprec <- read.csv(here("data/rwi_climate", "rprec_selected.csv"), header = TRUE, sep = ',') %>% mutate(v_name = 'Prec')

@@

label_site <- c("SJ" = "SJ",
                "caH" = "CA-High",
                "caL" = "CA-Low")

theme_ring_climate <- list(
  geom_hline(yintercept = 0, colour="gray"), 
  geom_bar(aes(fill=site), stat="identity", position = position_dodge2(padding = 0.1)),
  geom_text(aes(label = sig, x = id, y = ifelse(coef > 0, (coef + 0.02), (coef - 0.06))), 
            position = position_dodge(.9), size = 6), 
  theme_bw() +
  theme(axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_line(size=.5),
        axis.text.x = element_text(angle = 90, vjust = 0.3)),
  scale_fill_manual(values = c("#B2DF8A","#1F78B4","#A6CEE3"), labels = label_site)
)

p_tmax <- df %>% filter(v_name == 'T max') %>% 
  ggplot(aes(x=id, y=coef, group = site)) + 
  theme_ring_climate + 
  scale_y_continuous(limits = c(-0.4, 0.4), 
                     breaks = c(-0.4, -0.2, 0, 0.2, 0.4)) +
  scale_x_continuous(breaks = unique(df$id), labels = unique(df$month_name), position = 'bottom') + 
  ylab('Correlation') + 
  annotate("text", x = 2, y = 0.35, label = "T max", fontface =2, size = 5)

p_tmin <- df %>% filter(v_name == 'T min') %>% 
  ggplot(aes(x=id, y=coef, group = site)) + 
  theme_ring_climate + 
  scale_y_continuous(limits = c(-0.4, 0.6), 
                     breaks = c(-0.4, -0.2, 0, 0.2, 0.4)) +
  scale_x_continuous(breaks = unique(df$id), labels = unique(df$month_name), position = 'top') + 
  ylab('Correlation') +
  annotate("text", x = 2, y = 0.35, label = "T min", fontface =2, size = 5)

# prec-spei

@@

p_prec_spei <- dfprec_spei %>% 
  ggplot(aes(x=id, y=coef, group = site)) + 
  geom_vline(xintercept=20) +
  theme_ring_climate + 
  scale_y_continuous(limits = c(-0.2, 0.8), 
                     breaks = c(-0.2, 0, 0.2, 0.4, 0.6, 0.8)) +
  scale_x_continuous(breaks = labelitas$id, labels = labelitas$month_name, position = 'top') + 
  ylab('Correlation') +
  annotate("text", x = 2, y = 0.75, label = "Prec", fontface =2, size = 5) + 
  annotate("text", x = 23, y = 0.75, label = "SPEI", fontface =2, size = 5)

ptemperatura <- p_tmin + theme(legend.position="none") + p_tmax + theme(legend.position="bottom") + plot_layout(ncol=1)

ring_climate <- p_prec_spei + theme(legend.position="none") + ptemperatura + plot_layout(ncol=1, heights = c(2,3), tag_level = 'keep')  + plot_annotation(tag_levels = c('a','b'))

ring_climate
```

>>>>

---
output: 
  word_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("ggrepel")
library("scales")

@@

#####

```{r data}
rtmax <- read.csv(here("data/rwi_climate", "rtmax_selected.csv"), header = TRUE, sep = ',') %>% mutate(v_name = 'T max')
rtmin <- read.csv(here("data/rwi_climate", "rtmin_selected.csv"), header = TRUE, sep = ',') %>% mutate(v_name = 'T min')
rsequia <- read.csv(here("data/rwi_climate", "rsequia_selected.csv"), header = TRUE, sep = ',') %>% mutate(v_name ='SPEI')
rprec <- read.csv(here("data/rwi_climate", "rprec_selected.csv"), header = TRUE, sep = ',') %>% mutate(v_name = 'Prec')

@@

label_site <- c("SJ" = "SJ",
                "caH" = "CA-High",
                "caL" = "CA-Low")
```

```{r plot_options}
theme_ring_climate <- list(
  geom_hline(yintercept = 0, colour="gray"), 
  geom_bar(aes(fill=site), stat="identity", position = position_dodge2(padding = 0.1)),
  geom_text(aes(label = sig, x = id, y = ifelse(coef > 0, (coef + 0.02), (coef - 0.06))), 
            position = position_dodge(.9), size = 6), 
  theme_bw() +
  theme(axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_line(size=.5),
        axis.text.x.top = element_text(angle = 90, hjust = 0, vjust = 0.5),
        axis.text.x.bottom = element_text(angle = 90, hjust = 1, vjust = 0.5)),
  scale_fill_manual(values = c("#B2DF8A","#1F78B4","#A6CEE3"), labels = label_site)
) 
```

```{r plot_tmax}
p_tmax <- df %>% filter(v_name == 'T max') %>% 
  ggplot(aes(x=id, y=coef, group = site)) + 
  theme_ring_climate + 
  scale_y_continuous(limits = c(-0.4, 0.4), 
                     breaks = c(-0.4, -0.2, 0, 0.2, 0.4)) +
  scale_x_continuous(breaks = unique(df$id), labels = unique(df$month_name), position = 'bottom') + 
  ylab('Correlation') + 
  annotate("text", x = 2, y = 0.35, label = "T max", fontface =2, size = 3.5)
```

```{r plot_tmin}
p_tmin <- df %>% filter(v_name == 'T min') %>% 
  ggplot(aes(x=id, y=coef, group = site)) + 
  theme_ring_climate + 
  scale_y_continuous(limits = c(-0.4, 0.4), 
                     breaks = c(-0.4, -0.2, 0, 0.2, 0.4)) +
  scale_x_continuous(breaks = unique(df$id), labels = unique(df$month_name), position = 'top') + 
  ylab('Correlation') +
  annotate("text", x = 2, y = 0.35, label = "T min", fontface =2, size = 3.5)
```

```{r plot_prec_spei}
# prec-spei

@@

p_prec_spei <- dfprec_spei %>% 
  ggplot(aes(x=id, y=coef, group = site)) + 
  geom_vline(xintercept=20) +
  theme_ring_climate + 
  scale_y_continuous(limits = c(-0.2, 0.8), 
                     breaks = c(-0.2, 0, 0.2, 0.4, 0.6, 0.8)) +
  scale_x_continuous(breaks = labelitas$id, labels = labelitas$month_name, position = 'top') + 
  ylab('Correlation') +
  annotate("text", x = 2, y = 0.75, label = "Prec", fontface =2, size = 3.5) + 
  annotate("text", x = 23, y = 0.75, label = "SPEI", fontface =2, size = 3.5)
```

```{r plot_all}
ptemperatura <- p_tmin + theme(legend.position="none") + p_tmax + theme(legend.position="bottom") + plot_layout(ncol=1)

ring_climate <- p_prec_spei + theme(legend.position="none") + ptemperatura + plot_layout(ncol=1, heights = c(2,3), tag_level = 'keep')  + plot_annotation(tag_levels = c('a','b')) & theme(plot.tag = element_text(size = 12, face = 2)) &
theme(plot.margin = margin(t=0,r=0,b=-2,l=0, "pt"))
```

**Figure 8.** Correlation coefficients obtained by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data: precipitation and 6-month SPEI (a), minimun (b) and maximun (c) temperatures. *green* bars: northern site (SJ); *light blue* bars: low-elevation southern site (CA-Low); and *dark blue* bars: high-elevation shouthern site (CA-High). Asteriks indicate significant ($P < 0.05$) correlation coefficients.

```{r plot_out, fig.height=7, dpi = 300, out.width = 0.8} 
ring_climate
```
